The variability of Betelgeuse explained by surface convection

### 2024-04-23 main.tex
Add files via upload

\section{Introduction}

Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods can be found in the 
literature usually clustered around the so-called long secondary period (LSP) of 2000 days, and other shorter periods of 400 and 200 days \citep{kiss_variability_2006}. The origin of the LSP remains poorly known, some tried to explain the LSP by the lifetime of the large convective cells present at the surface of RSG (e.g \cite{stothers_giant_2010}) or by the magnetic field \citep{wood_long_2004}. The shorter periods of 400 and 200 days seem to have a different origin than the LSP. By studying a large sample of RSG, \cite{kiss_variability_2006} attributed the 400 days period to radial pulsation modes, the 200 days period being attributed to the first overtone \citep{joyce_standing_2020}, which was in agreement with the historical work of \cite{stothers_pulsation_1969}.\\

Besides radial pulsation, surface activity leading to random brightness variation was also mentioned to explain the variability of Betelgeuse, being firstly theorised by \cite{schwarzschild_scale_1975}. Betelgeuse is known to present large convective cells with lifetimes in the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes 
in the span of one week. Bright convection cells near disk center will increase the integrated brightness of the star when 
compared with other situations where such cells are found near the edges. So even without calling for the formation of 
dust or other changes in opacity, one may argue that the simple evolution of convective patterns of the star may create a variability. Such 
variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
convective patterns. \\

@@

The other period found in the linear polarization and stokes $U$, the 330 days periods, is more or less close to the 400 days period often present in the literature (e.g \cite{kiss_variability_2006} found a period $388 \pm 30$ days). It has been shown by \cite{lopez_ariste_convective_2018} that large convective cells can live up to one year, thus this period would be associated to the timescale on which the largest granules evolve. This typical timescale has also been observed in the numerical simulations and might play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}. Once again, the 330 days periodicity being more present in stokes $U$ than $Q$ unravels a stochastic process that is due to convective motions. \

@@

The main observation  presented in this work is that the same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
spectra measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes.
%But such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
%amplitude at selected wavelengths but no others. % Pulsation appears to offer no coherent explanation to these observations, are hardly to the 
%observed changes in the variability since the great dimming.  The concurrence of observational facts, actual period values, evolution before 
%and after the dimming, the larger amplitude of the period peaks in polarizarion data, points towards a common origin for those variabilities. 
%And such origin cannot be pulsations.
The interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
in the light curve and in the spectropolarimetric observations. This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 days periodicyt to convective dynamics while the 400 days periodicity is associated to both stochastic pulsations and convective timescales of the largest granules.%the observed periods are related 
%to convective dynamics and not to radial pulsations of the star.

>>>>

\section{Introduction}
Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods can be found in the 
literature usually clustered around the so-called long secondary period (LSP) of 2000 days, and other shorter periods of 400 and 200 days \citep{kiss_variability_2006}. The origin of the LSP remains poorly known, some tried to explain the LSP by the lifetime of the large convective cells present at the surface of RSG (e.g \cite{stothers_giant_2010}) or by the magnetic field \citep{wood_long_2004}. The shorter periods of 400 and 200 days seem to have a different origin than the LSP. By studying a large sample of RSG, \cite{kiss_variability_2006} attributed the 400 days period to radial pulsation modes, the 200 days period being attributed to the first overtone \citep{joyce_standing_2020}, which was in agreement with the historical work of \cite{stothers_pulsation_1969}.\\

Besides radial pulsation, surface activity leading to random brightness variation was also mentioned to explain the variability of Betelgeuse, being firstly theorised by \cite{schwarzschild_scale_1975} and \cite{gray_mass_2008} proposed that the 400 days period was a consequence of the convective activity. Betelgeuse is known to present large convective cells with lifetimes in the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes 
in the span of one week. Bright convection cells near disk center will increase the integrated brightness of the star when 
compared with other situations where such cells are found near the edges. So even without calling for the formation of 
dust or other changes in opacity, one may argue that the simple evolution of convective patterns of the star may create a variability. Such 
variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
convective patterns \citep{gray_mass_2008}. \\

@@

The other period found in the linear polarization and stokes $U$, the 330 days periods, is more or less close to the 400 days period often present in the literature (e.g \cite{kiss_variability_2006} found a period $388 \pm 30$ days). It has been shown by \cite{lopez_ariste_convective_2018} that large convective cells can live up to one year, thus this period would be associated to the timescale on which the largest granules evolve. This typical timescale has also been observed in the numerical simulations and might play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}. Once again, the 330 days periodicity being more present in stokes $U$ than $Q$ unravels a stochastic process that is due to convective motions, and is consistent with the work of \cite{gray_mass_2008}, who attributed the 400 days period to the large convective cells.  \

@@

The main observation  presented in this work is that the same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
spectra measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes.
%But such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
%amplitude at selected wavelengths but no others. % Pulsation appears to offer no coherent explanation to these observations, are hardly to the 
%observed changes in the variability since the great dimming.  The concurrence of observational facts, actual period values, evolution before 
%and after the dimming, the larger amplitude of the period peaks in polarizarion data, points towards a common origin for those variabilities. 
%And such origin cannot be pulsations.
The interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
in the light curve and in the spectropolarimetric observations. This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 days periodicyt to convective dynamics while the 400 days periodicity is associated to both stochastic pulsations and convective timescales of the largest granules. This conclusion is in agreement with the scenario proposed by \cite{gray_mass_2008}, who attributed the variability of Betelgeuse to surface convection. 
%the observed periods are related 
%to convective dynamics and not to radial pulsations of the star.
